---
title: "Inference 2.0"
author: "Mr. Adams"
date: "2/4/2020"
output: github_document
---
## Load Libraries 
```{r setup, load libraries}
library(tidyverse)
library(dslabs)
```

## 1. Get the data

The name of the dataset is NAME_OF_DATASET. 
Add in the name of the dataset and run the code to view the dataset. 
```{r}
view(NAME_OF_DATASET)
```

What are the two variables in this dataset?

1. 
2. 

## 2. Examine the data

What variable do I want to study? 

What type of variable is it?

## 3. Go down the Quantitative Path or the Categorical Path

##Quantitatve Path 
If you selected a QUANTITATIVE VARIABLE use this section, if not, skip below to find info on how to work with categorical variables.

```{r Histogram for quant variable}

NAME_OF_DATASET %>%
  ggplot(aes(x = QUANTITATIVE_VARIABLE)) +
  geom_histogram()

```

What is the shape of the distribution of reported NAME_OF_DATASET in your sample?


```{r Summary stats for quant variable}
QUANTITATIVE_VARIABLE_sum_stats <- NAME_OF_DATASET %>%
    summarize(avg = mean(QUANTITATIVE_VARIABLE), med = median(QUANTITATIVE_VARIABLE), 
              standard_dev = sd(QUANTITATIVE_VARIABLE), sample_size = n(), iqr = IQR(QUANTITATIVE_VARIABLE))

QUANTITATIVE_VARIABLE_sum_stats
```

```{r Confidence interval for a mean}

t.test(NAME_OF_DATASET$QUANTITATIVE_VARIABLE, conf.level = 0.95)

```

Write a sentence intepreting the confidence interval you created.


## Categorical Path - 
If you selected a CATEGORICAL VARIABLE use this section to create a bar graph, generate stats, and calculate confidence Interval

```{r Bar graph for categorical variable}

NAME_OF_DATASET %>%
  ggplot(aes(x = CATEGORICAL_VARIABLE)) +
  geom_bar()
```

```{r Summary stats for categorical variable }

CATEGORICAL_VARIABLE_count <- NAME_OF_DATASET %>%
            group_by(CATEGORICAL_VARIABLE) %>%
            summarize(count = n()) %>%
            mutate(prop = count/sum(count)) %>%
            arrange(desc(count))
CATEGORICAL_VARIABLE_count

n <- sum(CATEGORICAL_VARIABLE_count$count)
n
```
```{r Confidence interval for proportion}
prop.test(NUMBE_OF_MALES, SAMPE_SIZE, conf.level = 0.95)
```

Write a sentence interpreting the confidence interval you created. 
