---
title: "Inference 2.0"
author: "Mr. Adams"
date: "2/4/2020"
output: github_document
---
## Load Libraries 
```{r setup, load libraries}
library(tidyverse)
library(dslabs)
library(gridExtra)
```

## 1. Get the data

The name of the dataset is heights. 
Run the following code to view the dataset. 
```{r}
view(heights)
```

What are the two variables in this dataset?

1. 
2. 

## 2. Examine the data

What variable do I want to study? 

What type of variable is it?

## 3. Go down the Quantitative Path or the Categorical Path

##Quantitatve Path 
If you selected a QUANTITATIVE VARIABLE use this section, if not, skip below to find info on how to work with categorical variables.

```{r Histogram for quant variable}

heights %>%
  ggplot(aes(x = height)) +
  geom_histogram()

```

What is the shape of the distribution of reported heights in your sample?


```{r Summary stats for quant variable}
height_sum_stats <- heights %>%
    summarize(avg = mean(height), med = median(height), 
              standard_dev = sd(height), sample_size = n(), iqr = IQR(height))

height_sum_stats
```

```{r Confidence interval for a mean}

t.test(heights$height, conf.level = 0.95)

```

Write a sentence intepreting the confidence interval you created.


## Categorical Path - 
If you selected a CATEGORICAL VARIABLE use this section to create a bar graph, generate stats, and calculate confidence Interval

```{r Bar graph for categorical variable}

heights %>%
  ggplot(aes(x = sex)) +
  geom_bar()
```

```{r Summary stats for categorical variable }

sex_count <- heights %>%
            group_by(sex) %>%
            summarize(count = n()) %>%
            mutate(prop = count/sum(count)) %>%
            arrange(desc(count))
sex_count

n <- sum(sex_count$count)
n
```

```{r Confidence interval for proportion}
prop.test(812, 1050, conf.level = 0.95)
```

Write a sentence interpreting the confidence interval you created. 

#Taking a random sample from a large dataset 
In this section you will take a random sample from a large dataset. Your sample will have a new name. Once you have that, you can jump back up to the top to create a confidence interval for either a mean or proportion. 

```{r sampling from large dataset}

View(diamonds)
set.seed(123)
diamonds_sample <- sample_n(diamonds, 1000)
View(diamonds_sample)

```

```{r}
diamonds_sample %>%
  ggplot(aes(x = color)) +
  geom_bar()

color_count <- diamonds_sample %>%
            group_by(color) %>%
            summarize(count = n()) %>%
            mutate(prop = count/sum(count)) %>%
            arrange(desc(count))
color_count

n <- sum(color_count$count)
n

prop.test(123, 1000, conf.level = 0.95)
```

# Filter and then construct Confidence Intervals for means

Create new datasets. 

```{r Create Datasets}

just_for_men <- heights %>%
                filter(sex == "Male")

just_for_women <- heights %>%
                filter(sex == "Female")

```

Visualize the distribtutions
You can put all on the same graph or create different ones for each category. 

```{r Visualize the distributions for each category}

heights %>%
  ggplot(aes(x = height, color = sex)) +
  geom_density()

heights %>%
  ggplot(aes(x = height)) +
  geom_histogram() +
  facet_wrap(~sex, nrow = 2)
```


Generate summary stats for each group

```{r Generate summary stats for each category}
heights %>%
  group_by(sex) %>%
  summarize(avg = mean(height), med = median(height), 
              standard_dev = sd(height), sample_size = n(), iqr = IQR(height))
```

Generate Confidence Intervals for each group

```{r Construct a confidence interval for each group}
t.test(just_for_men$height, conf.level = 0.95)

t.test(just_for_women$height, conf.level = 0.95)
```


