---
title: "Inference 2.0"
author: "Mr. Adams"
date: "2/4/2020"
output: github_document
---
## Load Libraries 
```{r setup, load libraries}
library(tidyverse)
library(dslabs)
```

## 1. Get the data

The name of the dataset is heights. 
Run the following code to view the dataset. 
```{r}
view(heights)
```

What are the two variables in this dataset?

1. 
2. 

## 2. Examine the data

What variable do I want to study? 

What type of variable is it?

## 3. Go down the Quantitative Path or the Categorical Path

##Quantitatve Path 
If you selected a QUANTITATIVE VARIABLE use this section, if not, skip below to find info on how to work with categorical variables.

```{r Histogram for quant variable}

heights %>%
  ggplot(aes(x = height)) +
  geom_histogram()

```

What is the shape of the distribution of reported heights in your sample?


```{r Summary stats for quant variable}
height_sum_stats <- heights %>%
    summarize(avg = mean(height), med = median(height), 
              standard_dev = sd(height), sample_size = n(), iqr = IQR(height))

height_sum_stats
```

```{r Confidence interval for a mean}

t.test(heights$height, conf.level = 0.95)

```

Write a sentence intepreting the confidence interval you created.


## Categorical Path - 
If you selected a CATEGORICAL VARIABLE use this section to create a bar graph, generate stats, and calculate confidence Interval

```{r Bar graph for categorical variable}

heights %>%
  ggplot(aes(x = sex)) +
  geom_bar()
```

```{r Summary stats for categorical variable }

sex_count <- heights %>%
            group_by(sex) %>%
            summarize(count = n()) %>%
            mutate(prop = count/sum(count)) %>%
            arrange(desc(count))
sex_count

n <- sum(sex_count$count)
n
```
```{r Confidence interval for proportion}
prop.test(812, 1050, conf.level = 0.95)
```

Write a sentence interpreting the confidence interval you created. 

