---
title: "Introduction to Visualizations for EDA"
author: "John Adams"
output: github_document
---

An essential aspect of exploratory data analysis is visualizing your data. It allows you to draw conclusions more easily. This document will walk through a few visualizations you can create to see the story your data tells. 

## Load Tidyverse Library
```{r tidyverse, warning=FALSE}
library(tidyverse)
library(readr)
```

## Load the Data

```{r get the data}
urlfile="https://raw.githubusercontent.com/jwadams25/high_school_Rstats/master/girls_bball_2018.csv"

RGVB_team <- read.csv(url(urlfile))
```

## Histogram

A histogram is a great way of seeing the distribution of quantitative data. Let's start by analyzing the distribution of the points scored by the Rivers Girls Varsity Basketball team in each of their games during the 2018-2019 season. The variable name in the data fram is Points_For 

```{r histogram distribution of points scored by Rivers}
points_hist <- RGVB_team %>%
            ggplot(aes(x = Points_For)) +
            geom_histogram()
points_hist
```

Now let's add some labels to your axes and give the histogram a title. To start, I copied and pasted the code from above. 

```{r histogram distribution of points for with labels}
points_hist <- RGVB_team %>%
            ggplot(aes(x = Points_For)) +
            geom_histogram() +
            labs(x = "Points Scored", y = "Number of Games", title = "Distribution of Points Scored Per Game By Rivers Girls Basketball", subtitle = "2018-2019 Season")
points_hist
```

You'll see that a message popped up that says in part "Pick a better value with 'binwidth'" Let's make that change by adding in bindwith = 10 Reference data visualization principles to make an appropriate binwidth for your analysis. 

```{r histogram distribution with new bins}
points_hist <- RGVB_team %>%
            ggplot(aes(x = Points_For)) +
            geom_histogram(binwidth = 5) +
            labs(x = "Points Scored", y = "Number of Games", title = "Distribution of Points Scored Per Game By Rivers Girls Basketball", subtitle = "2018-2019 Season")
points_hist
```

Now let's change the color and outline of the bars. 

```{r histogram distribution with style changes}
points_hist <- RGVB_team %>%
            ggplot(aes(x = Points_For)) +
            geom_histogram(binwidth = 5, color = "black", fill = "red") +
            labs(x = "Points Scored", y = "Number of Games", title = "Distribution of Points Scored Per Game By Rivers Girls Basketball", subtitle = "2018-2019 Season")
points_hist
```

To finish let's set the scales of the x and y axes, add in a line for the median, and change the theme to clean things up a bit. 
```{r histogram distribution and median line}

points_hist <- RGVB_team %>%
            ggplot(aes(x = Points_For)) +
            geom_histogram(binwidth = 5, color = "white", fill = "red") +
            labs(x = "Points Scored", y = "Number of Games", 
                 title = "Distribution of Points Scored Per Game By Rivers Girls Basketball", 
                 subtitle = "2018-2019 Season")  +
            scale_x_continuous(limits = c(0, 90),
                     breaks = seq(0, 90, 10))+
            scale_y_continuous(limits = c(0, 8),
                     breaks = seq(0, 8, 1)) +
        geom_vline(aes(xintercept = median(Points_For)), color = "black")  +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
            
points_hist
            
```

## Density Curve

There are times when a density curve will more clearly show the distribution of quantitative data. Let's say you are going through your EDA and create a histogram first and realize a density curve would be a better visualization:
- copy and paste your code
- change the name
- change geom_histogram to geom_density
- delete binwidth =
- take out the scales code for the y-axis 
- change the y-axis label to y = "".

```{r Density Curves for distribution}
points_density <- RGVB_team %>%
            ggplot(aes(x = Points_For)) +
            geom_density(color = "white", fill = "red") +
            labs(x = "Points Scored", y = "", 
                 title = "Distribution of Points Scored Per Game By Rivers Girls Basketball", 
                 subtitle = "2018-2019 Season")  +
            scale_x_continuous(limits = c(0, 90),
                     breaks = seq(0, 90, 10)) +
        geom_vline(aes(xintercept = median(Points_For)), color = "black")  +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
            
points_density 
```

## Box Plots

Box plots are great for comparing distributions. Let's use them to compare the points scored during home games vs away games. 

```{r Box Plots to Compare Distributions}
points_HVA_boxes <- RGVB_team %>%
            ggplot(aes(x = H_V_A , y = Points_For)) +
            geom_boxplot() +
            theme_classic()
points_HVA_boxes
```

Before adding any style changes to this visualization, we want to make it easier for our reader to make conclusions. Therefore, we want to reorder these boxes from the greatest median to the least. To see this change, look at how code after aes changed. 

I also added in some code to change the style. Try to figure out what each addition changed.

```{r Box Plots to Compare Distributions reorder}
points_HVA_boxes <- RGVB_team %>%
            ggplot(aes(x = reorder(H_V_A, Points_For, .desc = TRUE), y = Points_For, color = H_V_A)) +
            geom_boxplot() +
            geom_jitter(alpha = 0.3) +
            scale_color_manual(values = c("black", "red")) +
            labs(title = "Distribution of Points Scored by Rivers Girls Basketball During Home and Away", subtitle = "2018-2019 Season", x = "", y = "Points Scored", color = "") +
            theme_classic()
points_HVA_boxes

```

Sometimes box plots are easier to read when they are stacked on top of one another. Use coord_flip to do that. Let's also add in labels while we are here. You will do that using the labs code like you did when making the histograms and density curves. 

```{r Box Plots to Compare Distributions add titles and flip}
flip_points_HVA_boxes <- RGVB_team %>%
            ggplot(aes(x = reorder(H_V_A, Points_For, .desc = TRUE), y = Points_For, color = H_V_A)) +
            geom_boxplot() +
            geom_jitter(alpha = 0.3) +
            scale_color_manual(values = c("black", "red")) +
            labs(title = "Distribution of Points Scored by Rivers Girls Basketball During Home and Away", subtitle = "2018-2019 Season", x = "", y = "Points Scored", color = "") +
            theme_classic() +
            coord_flip()
flip_points_HVA_boxes
```

## Bar Plots 

We'll use the bar plot to visualize categorical data. In this first example, we will look at the marginal distribution of result. In other words, the heights of the bars will represent the number of games they won and lost. 

```{r Bar plot basic}

wl_bars <- RGVB_team %>%
          ggplot(aes(x = Result)) +
          geom_bar()
wl_bars
```

Because we would most likely want to highlight the largest bar, let's put that one first. To do this, we have to change the order of the bars to go from greatest to least. 

Let's start by making a table using the skills we learned in previous activities to see the number of games for each result. After, we will use that data frame to make our bar plot. 

In this example, I have not made changes to the style. Knowing how style changes were made to previous visualizations, try to make similar changes to this one. If you are confused, start by changing the title and axes labels. 

```{r Bar plot ordering bars}

wl_count <- RGVB_team %>%
            group_by(Result) %>%
            summarize(count = n()) %>%
            arrange(desc(count))
wl_count

wl_bar_order <- wl_count %>%
          ggplot(aes(x = reorder(Result, -count), y = count)) +
          geom_bar(stat = 'identity')
wl_bar_order
```

## 100% Stacked Bar Graph

When you want to examine conditional distributions, a good option is a 100% stacked bar graph. For this example, we want to look at the proportion of home games and away games were won i.e., You will be able to answer the question, "Did they win a higher percentage of games at home or away?"
It's important to remember that you use this when studying two categorical variables.

```{r 100 stacked}
wl_venue <- RGVB_team %>%
            ggplot(aes(H_V_A, fill = Result)) +
            geom_bar(position = "fill") +
            labs(x = "", y = "Proportion", fill = "") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
wl_venue
```
